### 2018-06-13 paper/5-implementations.tex
Work on S5

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Approximation of task dependencies: a map of known upper bounds of dependencies
type ApproximateDependencies k = Map k [k]
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Build information: dirty keys, dependency approximation, and calculation chain
type CalcChain k = [k]
type ExcelInfo k = ((Set k, ApproximateDependencies k), CalcChain k)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Excel build system
excel :: Ord k => Build Monad (ExcelInfo k) k v
excel = restarting approximateRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Restarting scheduler
restarting :: Ord k => Rebuilder Monad i k v -> Build Monad (i, Chain k) k v
restarting rebuilder tasks key = execState $ do
    chain    <- gets (snd . getInfo)
    newChain <- go Set.empty $ chain ++ [key | key `notElem` chain]
    modify . mapInfo $ \(i, _) -> (i, newChain)
  where
    go :: Set k -> Chain k -> State (Store (i, Chain k) k v) (Chain k)
    go _    []     = return []
    go done (k:ks) = case tasks k of
        Nothing -> (k :) <$> go (Set.insert k done) ks
        Just task -> do
            store <- get
            let value = getValue k store
                newTask :: Task (MonadState i) k (Either k v)
                newTask = try $ rebuilder k value task -- `try' is defined below
                newFetch :: k -> State i (Either k v)
                newFetch k | k `Set.member` done = return $ Right (getValue k store)
                           | otherwise           = return $ Left k
            case runState (run newTask newFetch) (fst $ getInfo store) of
                (Left dep, _) -> go done $ [ dep | dep `notElem` ks ] ++ ks ++ [k]
                (Right newValue, newInfo) -> do
                    modify $ putInfo (newInfo, []) . updateValue k value newValue
                    (k :) <$> go (Set.insert k done) ks
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Convert a total task into a task accepting a partial fetch callback
try :: Task (MonadState i) k v -> Task (MonadState i) k (Either e v)
try task = Task $ \fetch -> runExceptT $ run task (ExceptT . fetch)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Rebuilder based on approximate dependencies
approximateRebuilder :: Ord k => Rebuilder Monad (Set k, ApproximateDependencies k) k v
approximateRebuilder key value task = Task $ \fetch -> do
    (isDirty, deps) <- get
    let dirty = key `Set.member` isDirty || case Map.lookup key deps of
                                                Nothing -> True
                                                Just ks -> any (`Set.member` isDirty) ks
    if not dirty
    then return value
    else do
        put (Set.insert key isDirty, deps)
        run task fetch
\end{minted}
\vspace{-2mm}
\caption{An implementation of \Excel using our framework.}\label{fig-excel-implementation}
\vspace{-2mm}
\end{figure}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Determine whether a trace is relevant to the current state
traceMatch :: (Monad m, Eq k)
    => (k -> Hash v -> m Bool) -> k -> [Trace k v] -> m [Hash v]
traceMatch check key ts = mapMaybeM f ts
    where f (Trace k dkv v) = do
                b <- return (key == k) &&^ allM (uncurry check) dkv
                return $ if b then Just v else Nothing
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Recursive dependency strategy
recursive :: Eq k => (k -> (k -> State (Store i k v, [k]) v)
                  -> State (Store i k v, [k]) (v, [k])
                  -> State (Store i k v, [k]) ())
                  -> Build Monad i k v
recursive process task key store = fst $ execState (ensure key) (store, [@@])
    where
        ensure key = do
            let fetch k = do ensure k; gets (getValue k . fst)
            done <- gets snd
            when (key `notElem` done) $ do
                modify $ \(s, done) -> (s, key:done)
                case track task fetch key of -- track is defined in @\S\ref{secdeps}@
                    Nothing -> return ()
                    Just act -> process key fetch act
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Shake build system
shake :: (Eq k, Hashable v) => Build Monad [Trace k v] k v
shake = recursive $ \key fetch act -> do
    traces <- gets (getInfo . fst)
    poss <- traceMatch (\k v -> (==) v . hash <$> fetch k) key traces
    current <- gets (getHash key . fst)
    when (current `notElem` poss) $ do
        (v, ds) <- act
        modify $ \(s, done) ->
            let t = Trace key [(d, getHash d s) | d <- ds] (getHash key s)
            in (putInfo (t : getInfo s) (putValue key v s), done)

\end{minted}
\vspace{-3mm}
\caption{An implementation of \Shake using our framework.}\label{fig-shake-implementation}
\vspace{1mm}
\end{figure}

>>>>

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Approximate task dependencies: a map of known upper bounds of dependencies
type ApproximateDependencies k = Map k [k]
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Build information: dirty keys, approximate dependencies, and calculation chain
type CalcChain k = [k]
type ExcelInfo k = ((Set k, ApproximateDependencies k), CalcChain k)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Excel build system
excel :: Ord k => Build Monad (ExcelInfo k) k v
excel = restarting approximateRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Restarting scheduler
restarting :: Ord k => Rebuilder Monad i k v -> Build Monad (i, Chain k) k v
restarting rebuilder tasks key = execState $ do
    chain    <- gets (snd . getInfo)
    newChain <- go Set.empty $ chain ++ [key | key `notElem` chain]
    modify . mapInfo $ \(i, _) -> (i, newChain)
  where
    go :: Set k -> Chain k -> State (Store (i, Chain k) k v) (Chain k)
    go _    []     = return []
    go done (k:ks) = case tasks k of
        Nothing -> (k :) <$> go (Set.insert k done) ks
        Just task -> do
            store <- get
            let value = getValue k store
                newTask :: Task (MonadState i) k (Either k v)
                newTask = try $ rebuilder k value task -- `try' is defined below
                newFetch :: k -> State i (Either k v)
                newFetch k | k `Set.member` done = return $ Right (getValue k store)
                           | otherwise           = return $ Left k
            case runState (run newTask newFetch) (fst $ getInfo store) of
                (Left dep, _) -> go done $ [ dep | dep `notElem` ks ] ++ ks ++ [k]
                (Right newValue, newInfo) -> do
                    modify $ putInfo (newInfo, []) . updateValue k value newValue
                    (k :) <$> go (Set.insert k done) ks
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Convert a total task into a task accepting a partial fetch callback
try :: Task (MonadState i) k v -> Task (MonadState i) k (Either e v)
try task = Task $ \fetch -> runExceptT $ run task (ExceptT . fetch)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Rebuilder based on approximate dependencies
approximateRebuilder :: Ord k => Rebuilder Monad (Set k, ApproximateDependencies k) k v
approximateRebuilder key value task = Task $ \fetch -> do
    (isDirty, deps) <- get
    let dirty = key `Set.member` isDirty || case Map.lookup key deps of
                                                Nothing -> True
                                                Just ks -> any (`Set.member` isDirty) ks
    if not dirty
    then return value
    else do
        put (Set.insert key isDirty, deps)
        run task fetch
\end{minted}
\vspace{-2mm}
\caption{An implementation of \Excel using our framework.}\label{fig-excel-implementation}
\vspace{-2mm}
\end{figure}

@@

\begin{figure}
\begin{minted}[fontsize=\small]{haskell}
-- Shake build system
shake :: (Ord k, Hashable v) => Build Monad (VT k v) k v
shake = recursive vtRebuilder
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Recursive dependency strategy
recursive :: Ord k => Rebuilder Monad i k v -> Build Monad i k v
recursive rebuilder tasks key store = fst $ execState (fetch key) (store, Set.empty)
  where
    fetch :: k -> State (Store i k v, Set k) v
    fetch key = case tasks key of
        Nothing -> gets (getValue key . fst)
        Just task -> do
            done <- gets snd
            when (key `Set.notMember` done) $ do
                value <- gets (getValue key . fst)
                let newTask :: Task (MonadState i) k v
                    newTask = rebuilder key value task
                    newFetch :: k -> StateT i (State (Store i k v, Set k)) v
                    newFetch k = do v <- lift $ fetch k
                                    put =<< lift (gets (getInfo . fst))
                                    return v
                info <- gets (getInfo . fst)
                (newValue, newInfo) <- runStateT (run newTask newFetch) info
                modify $ \(s, _) ->
                    (putInfo newInfo $ putValue key newValue s, Set.insert key done)
            gets (getValue key . fst)
\end{minted}
\vspace{1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Rebuilder based on verifying traces
vtRebuilder :: (Eq k, Hashable v) => Rebuilder Monad (VT k v) k v
vtRebuilder key value task = Task $ \fetch -> do
    vt <- get
    dirty <- not <$> verifyVT key value (fmap hash . fetch) vt
    if not dirty
    then return value
    else do
        (newValue, deps) <- trackM task fetch
        put =<< recordVT key newValue deps (fmap hash . fetch) =<< get
        return newValue
\end{minted}
\vspace{-3mm}
\caption{An implementation of \Shake using our framework.}\label{fig-shake-implementation}
\vspace{1mm}
\end{figure}
